---
title: "p8105_hw1_tq2171"
author: "Tingyu Qian"
date: "2024-09-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 1

## load the dataset
```{r}
# Load the 'penguins' dataset from the 'palmerpenguins' package
data("penguins", package = "palmerpenguins")

# Get the number of rows in the 'penguins' dataset
nrow(penguins)

# Get the number of columns in the 'penguins' dataset
ncol(penguins)

# Display the structure of the 'penguins' dataset, including data types and first few entries
str(penguins)

# Calculate the mean of the 'flipper_length_mm' column, excluding any missing values (NA)
mean(penguins$flipper_length_mm, na.rm = TRUE)
```

This is a dataset called "penguins", it contains 344 observations and 8 variables, which means there are 344 rows and 8 columns. Name and the type of the 8 variables are the following: species (factor), island (factor), bill_length_mm (numeric), bill_depth_mm (numeric), flipper_length_mm (integer), body_mass_g (integer), sex (factor) and the year (integer). Due to the variable flipper_length_mm contains missing value, so when we calculate the mean of flipper_length_mm, we need to drop the missing value, whcih use the na.rm=TRUE, and after that, we can know the mean of flipper length is 200.9152.
