---
title: "p8105_hw1_tq2171"
author: "Tingyu Qian"
date: "2024-09-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load the dataset
```{r}
data("penguins", package = "palmerpenguins")
nrow(penguins)
ncol(penguins)
str(penguins)
mean(penguins$flipper_length_mm, na.rm = TRUE)
```
This is a dataset called "penguins", it contains 344 observations and 8 variables, which means there are 344 rows and 8 columns. Name and the type of the 8 variables are the following: species (factor), island (factor), bill_length_mm (numeric), bill_depth_mm (numeric), flipper_length_mm (integer), body_mass_g (integer), sex (factor) and the year (integer). Due to the variable flipper_length_mm contains missing value, so when we calculate the mean of flipper_length_mm, we need to drop the missing value, whcih use the na.rm=TRUE, and after that, we can know the mean of flipper length is 200.9152.

## make the scatterplot
```{r}
# Load the ggplot2 library
library(ggplot2)

# Create a scatterplot
scatter_plot<-
  ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm, colour = species)) +
  geom_point() +
  labs(title = "Scatterplot of Bill Length vs Flipper Length", x = "bill_length_mm", y = "flipper_length_mm") +
  theme_minimal()

# Save the plot to the project directory
ggsave("scatterplot.png", plot = scatter_plot, width = 6, height = 4, dpi = 300)
```

