---
title: "p8105_hw1_tq2171"
author: "Tingyu Qian"
date: "2024-09-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 1

## load the dataset
```{r}
# Load the 'penguins' dataset from the 'palmerpenguins' package
data("penguins", package = "palmerpenguins")

# Get the number of rows in the 'penguins' dataset
nrow(penguins)

# Get the number of columns in the 'penguins' dataset
ncol(penguins)

# Display the structure of the 'penguins' dataset, including data types and first few entries
str(penguins)

# Calculate the mean of the 'flipper_length_mm' column, excluding any missing values (NA)
mean(penguins$flipper_length_mm, na.rm = TRUE)
```

This is a dataset called "penguins", it contains 344 observations and 8 variables, which means there are 344 rows and 8 columns. Name and the type of the 8 variables are the following: species (factor), island (factor), bill_length_mm (numeric), bill_depth_mm (numeric), flipper_length_mm (integer), body_mass_g (integer), sex (factor) and the year (integer). Due to the variable flipper_length_mm contains missing value, so when we calculate the mean of flipper_length_mm, we need to drop the missing value, whcih use the na.rm=TRUE, and after that, we can know the mean of flipper length is 200.9152.


## make the scatterplot
```{r}
# Load the ggplot2 library
library(ggplot2)

# Create a scatterplot
scatter_plot<-
  ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm, colour = species)) +
  geom_point() +
  labs(title = "Scatterplot of Bill Length vs Flipper Length", x = "bill_length_mm", y = "flipper_length_mm") +
  theme_minimal()

# Save the plot to the project directory
ggsave("scatterplot.png", plot = scatter_plot, width = 6, height = 4, dpi = 300)
```

# Homework 2
```{r}
# Load the tidyverse package
library(tidyverse)

# Set a seed for reproducibility
set.seed(123)

# Generate a random sample from a standard normal distribution
random_sample <- rnorm(10)

# Create the data frame with the corrected logical vector
df <- data.frame(
  random_sample = random_sample, # Random sample from standard Normal distribution
  logical_vector = random_sample > 0, # Logical vector indicating if each element of the sample > 0
  character_vector = sample(letters[1:10], 10, replace = TRUE), # Character vector of length 10
  factor_vector = factor(sample(c("Level1", "Level2", "Level3"), 10, replace = TRUE)) # Factor vector with 3 levels
)

print(df)
```

```{r}
# Mean of the numeric random sample
mean(df %>% pull(random_sample)) # This will work

# Mean of the logical vector
mean(df %>% pull(logical_vector)) # This will work (logical values are coerced to numeric: TRUE = 1, FALSE = 0)

# Mean of the character vector
mean(df %>% pull(character_vector)) # This will not work (character cannot be coerced to numeric)

# Mean of the factor vector
mean(df %>% pull(factor_vector)) # This will not work (factors are internally stored as integers but have a level attribute)
```
The mean calculation works for the numeric random sample and the logical vector but fails for the character and factor vectors due to how R handles different data types. The mean of the numeric random sample works because the values are inherently numeric, allowing for straightforward mathematical operations. Similarly, the mean of the logical vector works because R automatically coerces logical values (`TRUE` and `FALSE`) into numeric values (`1` and `0`), respectively, making it possible to compute the mean as the proportion of `TRUE` values. However, the mean calculation does not work for the character vector because characters are text data that cannot be directly converted into numbers; R does not know how to interpret or average text strings. For the factor vector, while it is stored internally as integers, it represents categorical data with associated levels. The levels have meaning as categories, not as numeric values, so taking their mean does not make sense and will not yield a meaningful result.

```{r}
# Coerce logical vector to numeric
numeric_logical <- as.numeric(df$logical_vector)

# Coerce character vector to numeric
numeric_character <- as.numeric(df$character_vector)

# Coerce factor vector to numeric
numeric_factor <- as.numeric(df$factor_vector)
```

The `as.numeric()` function in R successfully converts logical vectors and factor vectors to numeric form, but it fails to convert character vectors because of how these data types are internally represented. Logical vectors consist of `TRUE` and `FALSE` values, which have straightforward numeric equivalents: `TRUE` is converted to `1` and `FALSE` to `0`. This simple binary nature allows `as.numeric()` to perform a direct conversion. Similarly, factor vectors are stored as integers under the hood, with each factor level corresponding to a unique integer code. When `as.numeric()` is applied to a factor, it retrieves these underlying integer codes, making the conversion to numeric possible. In contrast, character vectors consist of text strings, which have no inherent numeric representation. R cannot assign a meaningful number to arbitrary text like "apple" or "dog," so `as.numeric()` fails to convert characters and instead returns `NA` with a warning. This difference highlights why `as.numeric()` works for logical and factor vectors but not for character vectors.

```{r}
# Calculate the mean of logic vector again
mean_numeric_logical <- mean(numeric_logical)  # This will work, calculating the proportion of TRUE values

# Calculate the mean of character vector again
mean_numeric_character <- mean(numeric_character, na.rm = TRUE)  # This will return NA or an error due to NA values

# Calculate the mean of factor vector again
mean_numeric_factor <- mean(numeric_factor)  # This will work, but the result may not be meaningful

# Print the results
print(mean(df %>% pull(random_sample))) 
print(mean_numeric_logical)    # Proportion of TRUE values
print(mean_numeric_character)  # Will be NA
print(mean_numeric_factor)     # Mean of factor levels' integer codes
```